---
title: "Exercise2"
author: "Kostas Kalogeropoulos"
date: "16/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Model and priors

Let $y=(y_1,\dots,y_n)$ be a r.s. from a N($\mu,v$) where $v$ is distributed according to an IGamma($\frac{N}{2},\frac{N}{2}\sigma^2$). The parameter $N$ is assumed to be known. Finalise the model with an improper prior on $\mu,\sigma^2$, $\pi(\mu,\sigma^2)\propto 1$.

### Part b: Gibbs Sampler details

The full conditional posterior for $v$ is
\begin{align*}
  \pi(v|\mu,\sigma^2,y)&  \propto v^{-\frac{n+N}{2}-1} \exp\left(-\frac{\frac{\sum_{i=1}^n(y_i-\mu)^2}{2}}{v}\right) \exp\left(-\frac{\frac{N}{2}\sigma^2}{ v}\right)\\
  &=v^{-\frac{n+N}{2}-1}  \exp\left(-\frac{\frac{N\sigma^2+\sum_{i=1}^n(y_i-\mu)^2}{2}}{v}\right)\\
  &\overset{\mathcal{D}}{=} \text{IGamma}\left(\frac{n+N}{2}, \frac{N\sigma^2+\sum_{i=1}^n(y_i-\mu)^2}{2}\right),
\end{align*}

whereas for $\mu$ is

\begin{align*}
  \pi(\mu|v,\sigma^2,y)&\propto \exp\left(-\sum_{i=1}^n\frac{(\mu^2-2\mu y_i+y_i^2)}{2v}\right)\propto\exp\left(-\frac{n\mu^2-2\mu \sum_{i=1}^n y_i}{2v} \right)\\
  &=\exp\left(-\frac{\mu^2-2\mu\bar{y}}{2v/n}\right)
  \overset{\mathcal{D}}{=} \text{ N} \left(\bar{y},\frac{v}{n}\right),
\end{align*}

and for $\sigma^2$

\begin{align*}
  \pi(\sigma^2|v,\mu,y)&\propto (\sigma^2)^{N/2 +1 -1} \exp\left(-\frac{N}{2v}\sigma^2\right)
  \overset{\mathcal{D}}{=} \text{ Gamma} \left(N/2 +1,\frac{N}{2v}\right),
\end{align*}


### Part c: Simulate Data and Run Gibbs Sampler

```{r cars}
n = 200 #sample size
mu_true = 2
v_true = 1
sigma2_true = 1
N = 20
y = rnorm(n,mu_true,sqrt(v_true))
ybar = mean(y)
S2 = sum((y-ybar)**2)
c(ybar,S2/(n-1))
dat=list(y=y,n=n,N=N)
```


```{r}
Niter = 10000 # number of iterations

#create object mus and sigma2s to store the Markov chain samples, 
#with $\pi(\theta,\sigma^2|y)$ as stationary distribution
out_mu = rep(0,Niter)
out_v = rep(0,Niter)
out_sigma2 = rep(0,Niter)


#Set initial values for mu and sigma2
mu = 10
sigma2 = 5
v = 4

#Main loop of Gibbs Sampler
for (iter in 1:Niter){
  # Store mu and sigma2 values
  out_mu[iter] = mu 
  out_v[iter] = v
  out_sigma2[iter] = sigma2
  
  #update mu given v sigma2 (and y)
  mu = rnorm(1,ybar,sqrt(v/n))
  
  #update v given mu sigma2 (and y)
  A = 0.5*(n + N)                  #a parameter of the full conditional of v
  B = N*sigma2 + sum((y-mu)**2)  #B parameter of the full conditional of sigma2
  v = 1/rgamma(1,A,B/2)
  
  #update sigma2 given mu (and y)
  sigma2  = rgamma(1,1+0.5*N, N/(2*v));
}
```

### Part d: Check MCMC output

```{r}
plot(out_mu,type='l')
plot(out_v,type='l')
plot(out_sigma2,type = 'l')
print('mu')
print(c(mean(out_mu),median(out_mu),quantile(out_mu,probs=c(0.025,0.975))))
print('v')
print(c(mean(out_v),median(out_v),quantile(out_v,probs=c(0.025,0.975))))
print('sigma2')
print(c(mean(out_sigma2),median(out_sigma2),quantile(out_sigma2,probs=c(0.025,0.975))))
```


### Part e: Repear using Stan


```{r}
library(rstan)
rstan_options(auto_write = TRUE)
```

```{r}
fit <- stan(file = 'Ex2e.stan', data = dat, init=0,chains=1,seed=1)
```

```{r}
print(fit)
plot(fit)
```