---
title: "ST308 Computer Workshop MCMC part 2"
output: html_document
---

## Demonstration

The dataset contains cases from a study that was conducted between, 1958 and 1970 at the University of Chicago's Billings Hospital on the survival of patients who had undergone surgery for breastcancer.

Four variables:
  * Age of patient at time of operation 
  * Patient's year of operation (year - 1900)
  * Number of positive axillary nodes detected 
  * Survival status (1 = the patient survived 5 years or longer, 2 = the patient died within 5 year)

First we load the data and fit a logistic regression model with MLE

```{r}
library(data.table)
url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/haberman/haberman.data"
biopsy <- fread(url, sep = ",", header = FALSE)
column.names <- c('age', 'year', 'nodes', 'survival')
setnames(biopsy, column.names)
biopsy$survival[biopsy$survival==2] = 0;
summary(biopsy)
fit1 = glm(survival~age+year+nodes,family=binomial,data=biopsy)
summary(fit1)
```

The task is to run a MCMC sampler via RStan to sample from the posterior of the logistic regression coefficients $\beta$. Normal priors can be assigned on the $\beta$ coefficients with zero mean and some reasonably large variance -say 100. The model can be written as

$$y_i\sim Bernoulli(\pi_i),\;\; i=1,...,n$$

$$\pi_i = \text{logit}^{-1}(X\beta),$$
$$\beta_0 \sim N(0,100I_4),$$
where $y_i$ denotes the survival status, $X$ is the design matrix with the first column consisting on 1's, second column being the age variable, 3rd column the year variable and 4th column the nodes detected variable for the $n$ individuals. $I_4$ denotes the identity matrix of dimension 4.


#### Code to fit MCMC with Stan

Load stan

```{r}
library(rstan)
rstan_options(auto_write = TRUE)
```

Prepare the data 

```{r}
y = biopsy$survival # y variable
n = length(y) # number of observations
X = cbind(rep(1,n),subset(biopsy, select = -survival)) # add one column of 1's for the constant and remove the y variable
X = cbind(rep(1,n),subset(biopsy, select = c(age,year,nodes))) # does the same as above
p = dim(X)[2] # number of variables plus the constant

biopsy_dat=list(n=n, p=p, X=X, y=y)
```

For the model see the file `LogisticRegression.stan`. Next we run MCMC

```{r}
fit <- stan(file = 'LogisticRegression.stan', data = biopsy_dat,chains=2,init=0,seed=1)
```

summarise posterior output

```{r}
print(fit)
traceplot(fit)
plot(fit, pars=c("beta[2]","beta[3]","beta[4]"),show_density = TRUE, ci_level = 0.95, fill_color = "green")
```

## Activity

Repeat but now fit a model with only the age and the nodes detected variables.

Put your code below

```{r}

```

