---
title: "ST308 - week 02"
output: html_document
---
      
In this workshop:

- Basic commands for vectors and matrices
- Data visualisation
- Loading data 

## Basic R Commands

**Vectors and sequences**

- Create vectors x and y
- Vector arithmetic


```{r setup, include=TRUE}

x <- c(1,3,2,5)
y <-  c(1,4,3,0)

length(x)
length(y)

x+y
2*x-y+1
```

The function `seq` creates a sequence of numbers

```{r}
#sequences
x <- seq(1,10)

#sequence with specified number of values
y <- seq(1,10,length.out=20)

```

`rm` is the function that removes objects from the environment

```{r}
rm(x,y)  #to remove objects x and y
#rm(list=ls())
```


**Matrices**

The function `matrix` creates a matrix object. 

To construct a 2X2 matrix with numbers from 1 to 4, do:


```{r}
?matrix
x=matrix(data=c(1,2,3,4), nrow=2, ncol=2)
x
x=matrix(c(1,2,3,4),2,2)

```

**Elementwise operations on matrices**

To find the square root and the square of each matrix's component:

```{r}
# Elementwise operations on matrices
sqrt(x)
x^2
```

****


## Indexing Data

Below are some commands for practicing in working with matrices and accessing their elements

```{r}
A=matrix(1:16,4,4)
A
A[2,3] 
A[c(1,3),c(2,4)]
A[1:3,2:4]
A[1:2,]
A[,1:2]
A[1,]
A[-c(1,3),]
A[-c(1,3),-c(1,3,4)]
dim(A)
```


## Drawing observations from normal distribution

```{r}

x=rnorm(50)

#for replicability set seed
set.seed(1303)
rnorm(50)

set.seed(3)
y=rnorm(100)
mean(y)
var(y)
sqrt(var(y))
sd(y)
```

**Correlations**

Recall the Pearson correlation measures the strength of linear relationshiop between 2 quantitative variables, and takes values [-1,1].

Task: Create 2 variables: 50 observations from $X \sim N(0,1)$, and $Y=X+N(50,0.1^2)$
Then compute the (Pearson's) correlation.

```{r}

y=x+rnorm(50,mean=50,sd=.1)
cor(x,y)
```


## Activity 1

1. Run the following chunk:

```{r}
A <- matrix(c(1,2,3,4),2,2,byrow=TRUE)
t(A) 
```

What does the function `t` do?

2. Draw 10 numbers from the normal distribution with mean 5 and variance 2. Store them in the vector x and calculate the sample mean and standard deviation. Put them in a 2 X 5 matrix by row and and raise each number to the power of three.

### Put your code below


## Data visualisation

R has a set of functions for creating plots. A popular package for data visualisation is ggplot. We will focus on the core R graphics.


**Scatterplot**

1st simulate data for 2 variables and store values in a data frame object.

```{r}
x1 <- rnorm(100)
y1 <- rnorm(100)

dat <- data.frame(x1,y1)
dat
```

`plot` produces a scatterplot

```{r}
plot(x1,y1) #note order matters

```

To save the plot as a pdf in your folder, do:

```{r}
pdf("Figure.pdf")
plot(x,y,col="blue")
dev.off()  #closes the plot and uploads it 
```


## Loading Data

The functions for loading data depends on the format of the data.

- `read.table` is used to load data that is stored as a txt or csv file. 
- `read.csv` is for csv

A common error when loading data is not checking whether or not the data comes with a header. 
Also, if you had not already done so at the beginning of your session, set the working directory to the folder where your files are stored.

**Example: Automobile data**

The data Auto.data.txt consists of variables on the profile of vehicles. The data set contains a header (check). 
Data sets may have variables with missing values. Here, `na.strings=?` sets anything with ? to NA.

Load the data and check all us ok using `head`:

```{r}
Auto <- read.table("Auto.txt", header=T,na.strings="?")
head(Auto)
```

`summary` gives the 5-point summary statistics  plus the mean for each variable.

```{r}
summary(Auto)
```

**Subsetting data**

How many vehicles have mpg>30? 
What is the distribution of mpg with horsepower = 90?

We can answer such question by using functions for subsetting the data. There is more than one way to subset data and it's a matter of preference.

The number of vehicles is:

```{r}
nrow(subset(Auto,mpg>10)) #or use dim
```

The histogram for mpg with horsepower =90 is:

```{r}
hist(Auto$mpg[Auto$horsepower==90]) # note the double =
```


## Activity 2

This exercise involves the Boston housing data set which is
part of the MASS library in R. You can read about it typing `?Boston` after loading the library. To load and view part of it use the following commands:

```{r}
library(MASS)
head(Boston)
```


(a) How many rows are in this data set? How many columns? What do the rows and columns represent?

(b) Are any of the predictors 'age', dis 'ptratio' appear to be associated with per capita crime rate? If so, explain the relationship.

(c) In this data set, how many of the suburbs average more than seven rooms per dwelling? More than eight rooms per dwelling?


### Put your code below

Knit the file and check the output!
