---
title: "Computer Class ST308 Week 10"
author: "Kostas Kalogeropoulos"
date: "26/03/2020"
output:
  html_document: default
  pdf_document: default
---

## Load data

```{r}
RadonData = read.csv('radon.csv',header=TRUE)
summary(RadonData)
```

convert the string (factor variable) county to numeric

```{r}
RadonData$county = as.numeric(RadonData$county)
```

Create a new binary variable y that takes the value 1 if the variable logRadon is above its median and 0 otherwise

```{r}
RadonData$y = RadonData$logRadon>median(RadonData$logRadon)
mean(RadonData$y)
```

## Model

Consider the following model

$$
 y_{ij} \sim \text{Bernoulli}[\pi_{ij}(X_{ij})],
$$ 
$$
 \log\left(\frac{\pi_{ij}(X_{ij})}{1-\pi_{ij}(X_{ij})}\right) = \alpha_{i} +\beta X_{ij},
$$
 $$
 \alpha_i \overset{\text{ind}}{\sim}N(\mu_{a},\sigma_{a}^2)
 $$
 $$
  \mu_{a}\sim N(0,10)
 $$
 $$
 \sigma_{a}\sim \text{HalfCauchy}(1)
 $$
 $$
 \beta \sim N(0,1000^2)
 $$


## Demonstration / Activity

Fit the above model in Stan and present the MCMC output. Put your code below.



Prepare the data for Stan by putting them in a list

```{r}
library(rstan)
radon_dat=list(y=RadonData$y, N=length(RadonData$logRadon), K=length(unique(RadonData$county)), x=RadonData$floor, county=RadonData$county)
```

Run Hamiltonian MCMC in Stan

```{r}
fit <- stan(file = 'RadonLogisticMultilevel.stan', data = radon_dat, init=0,chains=1,iter=4000,seed=1)
```

Analyse and interpret MCMC output

```{r}
print(fit,pars=c("mu_a","sigma_a","beta"))
plot(fit,pars=c("alpha"))
traceplot(fit,pars=c("mu_a","sigma_a","beta","alpha[49]","alpha[85]"))
```
