---
title: "ST308-week06"
output: html_document
---


You will need to load the core library for the course textbook:
```{r}
library(ISLR)
```


# Demonstration 

In this exercise, we will predict the number of applications received using the other variables in the `College` data set.

(a) Split the data set into a training set and a test set.

**Load and split the `College` data.**

```{r}
library(ISLR)
set.seed(11)
sum(is.na(College))
train.size <-  nrow(College) / 2
train <-  sample(1:nrow(College), train.size)
test <-  -train
College.train <-  College[train, ]
College.test <-  College[test, ]
```

(b) Fit a linear model using least squares on the training set, and
report the test error obtained. Select the model based on either the 

**Number of applications is the Apps variable.**
```{r}
lm.fit <-  lm(Apps ~ . , data = College.train)
lm.pred <-  predict(lm.fit, College.test)
mean((College.test[, "Apps"] - lm.pred)^2)
```

**Test RSS is 1538442**

Check the regression coefficients

```{r}
summary(lm.fit)
```
(c) Fit a ridge regression model on the training set, with $\lambda$ chosen by cross-validation. Report the test error obtained.

**Pick $\lambda$ using College.train and report error on College.test**

```{r}
library(glmnet)
set.seed(1)
train.mat <-  model.matrix(Apps ~ . , data = College.train)
test.mat <-  model.matrix(Apps ~ . , data = College.test)
grid <-  10 ^ seq(4, -2, length = 100)
mod.ridge <-  cv.glmnet(train.mat, College.train[, "Apps"], 
                        alpha = 0, lambda = grid, thresh = 1e-12)
lambda.best <-  mod.ridge$lambda.min
lambda.best
ridge.pred <-  predict(mod.ridge, newx = test.mat, s = lambda.best)
mean((College.test[, "Apps"] - ridge.pred)^2)
```
**Test RSS is slightly higher than OLS, 1608859.**

Check the ridge regression coefficients

```{r}
mod.ridge <-  glmnet(model.matrix(Apps ~ . , data = College), 
                     College[, "Apps"], alpha = 1,lambda=lambda.best)
predict(mod.ridge, s = lambda.best, type = "coefficients")
```

(d) Fit a lasso model on the training set, with $\lambda$ chosen by cross-validation. Report the test error obtained, along with the number of non-zero coefficient estimates.

**Pick $\lambda$ using College.train and report error on College.test.**
```{r}
set.seed(1)
mod.lasso <-  cv.glmnet(train.mat, College.train[, "Apps"], 
                        alpha = 1, lambda = grid, thresh = 1e-12)
lambda.best <-  mod.lasso$lambda.min
lambda.best
lasso.pred <-  predict(mod.lasso, newx = test.mat, s = lambda.best)
mean((College.test[, "Apps"] - lasso.pred)^2)
```
**Again, Test RSS is slightly higher than OLS, 1630061.**

**The Lasso coefficients look like**
```{r}
mod.lasso <-  glmnet(model.matrix(Apps ~ . , data = College), 
                     College[, "Apps"], alpha = 1)
predict(mod.lasso, s = lambda.best, type = "coefficients")
```


## Activity

We will now try to predict per capita crime rate in the `Boston` data set.


```{r}
set.seed(1)
library(MASS)
train.size <-  nrow(Boston) / 2
train <-  sample(1:nrow(Boston), train.size)
test <-  -train
Boston.train <-  Boston[train, ]
Boston.test <-  Boston[test, ]
lm.fit <-  lm(medv ~ . , data = Boston.train)
lm.pred <-  predict(lm.fit, Boston.test)
mean((Boston.test[, "medv"] - lm.pred)^2)
summary(lm.fit)
```

Repeat the previous tasks

**Ridge Regression**

```{r}
set.seed(2)
train <-  model.matrix(medv ~ . , data = Boston.train)
test <-  model.matrix(medv ~ . , data = Boston.test)
grid <-  10 ^ seq(4, -2, length = 100)
mod.ridge <-  cv.glmnet(train, Boston.train[, "medv"], 
                        alpha = 0, lambda = grid, thresh = 1e-12)
lambda.best <-  mod.ridge$lambda.min
lambda.best
ridge.pred <-  predict(mod.ridge, newx = test, s = lambda.best)
mean((Boston.test[, "medv"] - ridge.pred)^2)
mod.ridge <-  glmnet(model.matrix(medv ~ . , data = Boston.train), 
                     Boston.train[, "medv"], alpha = 1,lambda=lambda.best)
predict(mod.ridge, s = lambda.best, type = "coefficients")
```

**Lasso**
```{r}
set.seed(3)
mod.lasso <-  cv.glmnet(train, Boston.train[, "medv"], 
                        alpha = 1, lambda = grid, thresh = 1e-12)
lambda.best <-  mod.lasso$lambda.min
lambda.best
lasso.pred <-  predict(mod.lasso, newx = test, s = lambda.best)
mean((Boston.test[, "medv"] - lasso.pred)^2)
mod.rlasso <-  glmnet(model.matrix(medv ~ . , data = Boston.train), 
                     Boston.train[, "medv"], alpha = 1,lambda=lambda.best)
predict(mod.lasso, s = lambda.best, type = "coefficients")
```
