---
title: "Markov Chain Illustration"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Staionary distribution Illustration for a Markov Chain


First, we will simulate 500 points from the Markov chain mentioned in class -AR(1) model

```{r}
N=5000
Xa=vector(N+1,mode="numeric") #create an empty vector of length N+1
Xa0=8 				
Xa[1]=Xa0			#Set the first entry of the vector to Xa0 

#loop for a simulation of the Markov Chain
for (i in 1:N){
Xa[i+1]=rnorm(1,0.5*Xa[i],1)
}

plot(Xa,type="l")
```


Now we will repeat this to simulate 500 points from the same Markov chain but at a different starting value

```{r}
Xb=vector(N+1,mode="numeric") #create an empty vector of length N+1
Xb0=-7 				
Xb[1]=Xb0			#Set the first entry of the vector to Xa0 

#loop for a simulation of the Markov Chain
for (i in 1:N){
Xb[i+1]=rnorm(1,0.5*Xb[i],1)
}
plot(Xb,type="l")
```

Finally we will draw independent samples from the stationary distribution which is known in this case (now always true) to be the N$(0,1.333)$ distribution. We will compare these three sets of draws with graphs and summaries

```{r}
Xc = rnorm(N,0,sqrt(1.333)) 
plot(Xa,type="l",ylim=c(-Xa0,Xa0),col="purple")
lines(Xb,col="red")
lines(Xc)

mean(Xa)
mean(Xb)
var(Xa)
var(Xb)

plot(density(Xa),xlab="X",col="purple",ylim=c(0,0.37),lwd=2)
lines(density(Xb),col="red",lwd=2)
X=seq(from=-4.5,to=4.5,by=0.01)
lines(X,dnorm(X,0,sqrt(1.333)),lwd=2)
legend(4,0.15,c("true","chain 1","chain 2"),lty=c(1,1,1),lwd=c(2,2,2),
col=c("black","red","purple"))
```

